Modelling low Mach number stellar hydrodynamics with MAESTROeX

### 2019-10-02 astronum_2019_MAESTROeX/astronum-maestroex.tex
finished writing summary

\maestroex~is currently able to model spherically symmetric systems, however it has no support for rotation. \MarginPar{I would say it has no support for rotation that breaks the spherical symmetry} We wish to model convection in the interiors of massive stars, and it is known that these stars often have non-negligible rotational frequencies and that this rotation can have have significant effect on mixing at convective shell boundaries. We are therefore currently exploring several possible ways of implementing rotation in \maestroex.

The simplest way to include rotation is to ignore the effects of the centrifugal force, and only include the Coriolis force. This was done in \maestro, and should be a reasonable approximation for slowly rotating systems. 
\MarginPar{In Zingale 2011, equation (2) has the Coriolis and centrifugal terms.  Include that equation here?}

@@

\MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
The latest generation of supercomputers to come online (e.g.\ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs). This trend is set to continue: supercomputers are moving to new architectures in order to increase performance rather than simply adding more and more CPUs. In order for HPC codes to be able to exploit these latest machines, it is therefore becoming increasingly necessary to port codes to run on GPUs.

@@

As mentioned, our work porting \maestroex~to run on GPUs is still ongoing, however in \ref{fig:gpu_speedup}, we show the speedup we have achieved so far for a number of individual functions that have been offloaded to GPU. The plot shows the average execution time of each function, recorded for the 3d \texttt{reacting\_bubble} problem on the OLCF Summit machine. Both tests were run on a single node, consisting of two IBM POWER9 processors (together providing 42 physical CPU cores) and six NVIDIA Volta GPUs. The CPU test used 42 MPI processes, each with 4 OpenMP threads, and the GPU test used 6 MPI processes, each process associated with a single GPU. Note that the plot's $x$-axis scale is logarithmic. As can be seen, large speedups in excess of 30-40$\times$ were achieved for several of the functions when run on the GPU. Once we have finished porting the remaining routines to GPU and further optimised our code, we hope to achieve this sort of speedup for the entire timestep.

@@

More future work: SDC, more science, etc.

>>>>

\maestroex~is currently able to model spherically symmetric systems, however it has no support for rotation which breaks the spherical symmetry. We wish to model convection in the interiors of massive stars, and it is known that these stars often have non-negligible rotational frequencies and that this rotation can have have significant effect on mixing at convective shell boundaries. We are therefore currently exploring several possible ways of implementing rotation in \maestroex.

As described in \cite{Zingale2011}, to incorporate rotation in our equation set, we add the Coriolis and centrifugal terms to the velocity evolution equation:
\begin{equation}
    \pd{\Ub}{t} = -\Ub\cdot\nabla\Ub - \frac{\beta_0}{\rho}\nabla\left(\frac{\pi}{\beta_0}\right) - \frac{\rho - \rho_0}{\rho}\bm{e}_r - 2\boldsymbol{\Omega} \times \Ub - \boldsymbol{\Omega} \times (\boldsymbol{\Omega} \times \bm{r}),
\end{equation}
where $\boldsymbol{\Omega}$ is the angular velocity, $\bm{F}_\text{Coriolis}=- 2\boldsymbol{\Omega} \times \Ub$ is the Coriolis force and $\bm{F}_\text{centrifugal}= - \boldsymbol{\Omega} \times (\boldsymbol{\Omega} \times \bm{r})$ the centrifugal force.

The simplest way to include rotation is to ignore the effects of the centrifugal force, and only include the Coriolis force. This was done in \maestro, and should be a reasonable approximation for slowly rotating systems. 
% \MarginPar{In Zingale 2011, equation (2) has the Coriolis and centrifugal terms.  Include that equation here?}

@@

% \MarginPar{I feel like someone who knows what they're talking about should write/seriously edit this section}
The latest generation of supercomputers to come online (e.g.~ OLCF Summit, soon NERSC Perlmutter and OLCF Frontier) rely heavily on GPUs in order to achieve high performance while keeping energy consumption to a minimum (GPUs consume considerably less power per flop than CPUs). This trend is set to continue: supercomputers are moving to new architectures in order to increase performance rather than simply adding more and more CPUs. In order for HPC codes to be able to exploit these latest machines, it is therefore becoming increasingly necessary to port codes to run on GPUs.

@@

As mentioned, our work porting \maestroex~to run on GPUs is still ongoing, however in Figure~\ref{fig:gpu_speedup} we show the speedup we have achieved so far for a number of individual functions that have been offloaded to GPU. The plot shows the average execution time of each function, recorded for the 3d \texttt{reacting\_bubble} problem on the OLCF Summit machine. Both tests were run on a single node, consisting of two IBM POWER9 processors (together providing 42 physical CPU cores) and six NVIDIA Volta GPUs. The CPU test used 42 MPI processes, each with 4 OpenMP threads, and the GPU test used 6 MPI processes, each process associated with a single GPU. Note that the plot's $x$-axis scale is logarithmic. As can be seen, large speedups in excess of 30-40$\times$ were achieved for several of the functions when run on the GPU. Once we have finished porting the remaining routines to GPU and further optimised our code, we hope to achieve this sort of speedup for the entire timestep.

@@

Work has recently begun on implementing the new time integration strategy, spectral deferred corrections (SDC) that has recently been implemented in \castro~\cite{castro:sdc}. This method eliminates the coupling error between source terms and hydrodynamics incurred in operator splitting techniques. It is hoped that this shall be particularly useful for our simulations of rotating massive stars, where we have found energetic reactions in the star's core to be challenging to model using our current methods.
